---
title: "PICO 2: Prepare Data for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

## Create folders if missing:
root_dir <- knitr::opts_knit$get("root.dir")
figs_dir <- paste0(root_dir, "/figures")
results_dir <- paste0(root_dir, "/results")
data_derived_dir <- paste0(root_dir, "/data_derived")
if(!dir.exists(figs_dir))  dir.create(figs_dir)
if(!dir.exists(results_dir)) dir.create(results_dir)
if(!dir.exists(data_derived_dir)) dir.create(data_derived_dir)
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(readxl)
library(data.table)
library(writexl)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

Check existence of file with raw data:
```{r}
RAWDATA <- "data_raw/raw_data_butyrate_meta_analysis.xlsx"
stopifnot(file.exists(RAWDATA))
```

Preliminary reading of data treating only empty cells as `NA` (cells containing "NA" are _not_ treated as `NA`):
```{r, results='hide'}
## readxl may fail if the workbook is open in the desktop Excel application; browser application works fine.
d0 <- quick_read_excel(RAWDATA, na = c(""), sheet="PICO 2", guess_max = 1e4)
setDT(d0)
## Preview/inspect dataset:
dim(d0)
str(d0)
head(d0)
tail(d0)
```


### Check for expected columns (=variables) in dataset

```{r}
expected_col_names <- 
  c("forest.plot", "include", "outcome.cat", "outcome.subcat", "direction", "outcome", 
    "sample.type", "o.units", "author", 
    "year", "trialID", "age.group", "population.status", "disease", "intervention.type", 
    "duration", "followup", "i.dosis", "i.units", "control", "c.dosis", 
    "c.units", "n.e", "b.event.e", "b.m.e", "b.sd.e", "b.se.e", "b.med.e", 
    "b.min.e", "b.max.e", "b.q1.e", "b.q3.e", "b.ci.low.e", "b.ci.high.e", 
    "eot.event.e", "eot.m.e", "eot.sd.e", "eot.se.e", "eot.med.e", 
    "eot.min.e", "eot.max.e", "eot.q1.e", "eot.q3.e", "eot.ci.low.e", 
    "eot.ci.high.e", "n.c", "b.event.c", "b.m.c", "b.sd.c", 
    "b.se.c", "b.med.c", "b.min.c", "b.max.c", "b.q1.c", "b.q3.c", 
    "b.ci.low.c", "b.ci.high.c", "eot.event.c", "eot.m.c", "eot.sd.c", 
    "eot.se.c", "eot.med.c", "eot.min.c", "eot.max.c", "eot.q1.c", 
    "eot.q3.c", "eot.ci.low.c", "eot.ci.high.c", "notes")

if(!all(expected_col_names %in% names(d0))) {
  missing_var <- setdiff(expected_col_names, names(d0))
  stop("Required variables in data not available: ", paste(missing_var, collapse = ", "))
}
if(length(new_variables <- setdiff(names(d0), expected_col_names))) 
  warning("New variables detected in data: ", paste(new_variables, collapse = ", "))
```


### Check for empty cells in dataset 

Check for empty cells in analytic columns, excluding `forest.plot` and free-text columns.

```{r}
## Only read read columns until free-text columns (notes/comments):
max_col <- min(grep("notes|Notes|comment|Comment", names(d0))) - 1
d <- d0[, 1:max_col]

## Columns in which to ignore missing cells:
col_ignore <- "forest.plot"
stopifnot(all(col_ignore %in% names(d)))

## Identify cells with missing values:
x <- sapply(d[, -col_ignore, with = FALSE], function(x) sum(is.na(x)))
x <- x[x > 0]

if(length(x) > 0) {
  # Get list of places with empty cells:
  nm <- setdiff(names(x), col_ignore)
  message("Missing cells in the following columns (row numbers):")
  print(lapply(setNames(nm, nm), function(n) which(is.na(d0[, get(n)]))+1) )
  warning("Missing cells are not allowed!")
}
```


## Read and check data

Read the data treating `NA` as missing:
```{r, results='hide'}
# Read file again including only real data-rows:
d0 <- quick_read_excel(RAWDATA, na = c("", "NA"), guess_max = 1e4,
                       sheet="PICO 2") 
                
setDT(d0)
d <- d0[, 1:max_col][, -col_ignore, with = FALSE]
dim(d)
head(d)
tail(d)
str(d, list.len=1e3)
```

Prepend a row id column:
```{r}
d[, row_id := 1:.N]
setcolorder(d, "row_id")
```


### Check that all data columns are numeric 

Data columns end in `.e` or `.c`:
```{r}
nm <- names(d)
data_cols <- nm[grep("\\.e$|\\.c$", nm)]
not_is_numeric <- names(which(!sapply(d[, ..data_cols], is.numeric)))
## Coerce non-numeric data-variables to numeric:
for(nam in not_is_numeric) { 
  tmp <- d[, get(nam)]
  tmp_num <- suppressWarnings(as.numeric(tmp))
  if(sum(is.na(tmp_num)) > sum(is.na(tmp))) 
    warning("NAs introduced when coercing '", nam, "' to numeric")
  d[, (nam) := as.numeric(get(nam))]
}
```


### Tabulate non-numeric variables 

```{r}
nms <- nm[!sapply(d, is.numeric)]
tables <- lapply(setNames(nms, nms), function(col) {
  totpct(d[, .N, by=col][order(-N)])
})
write_xlsx(tables, path="results/PICO2_categorical_variable_tables.xlsx")
```


### Check consistency in reporting of mean/median/sd/se

Check that if mean/median is reported in the experimental group, it is also reported in the control group. 
```{r}
tmp <- d
stopifnot(
  tmp[is.na(eot.m.e) & !is.na(eot.m.c) | !is.na(eot.m.e) & is.na(eot.m.c), .N == 0],
  tmp[is.na(b.m.e) & !is.na(b.m.c) | !is.na(b.m.e) & is.na(b.m.c), .N == 0], 
  
  tmp[is.na(eot.med.e) & !is.na(eot.med.c) | !is.na(eot.med.e) & is.na(eot.med.c), .N == 0],
  tmp[is.na(b.med.e) & !is.na(b.med.c) | !is.na(b.med.e) & is.na(b.med.c), .N == 0], 
  
  tmp[is.na(eot.sd.e) & !is.na(eot.sd.c) | !is.na(eot.sd.e) & is.na(eot.sd.c), .N == 0],
  tmp[is.na(b.sd.e) & !is.na(b.sd.c) | !is.na(b.sd.e) & is.na(b.sd.c), .N == 0],
  
  tmp[is.na(eot.se.e) & !is.na(eot.se.c) | !is.na(eot.se.e) & is.na(eot.se.c), .N == 0],
  tmp[is.na(b.se.e) & !is.na(b.se.c) | !is.na(b.se.e) & is.na(b.se.c), .N == 0] 
)

```


### Check consistency of min-max and low-high variables

Check that if `q1` is non-NA, then `q3` is also non-NA and vice versa. Also check min-max and lower-upper confidence limits as well.
```{r}
stopifnot(
  ## Check Q1 and Q3 data:
  d[(is.na(eot.q1.e) & !is.na(eot.q3.e)) |  (!is.na(eot.q1.e) & is.na(eot.q3.e)), .N == 0],
  d[(is.na(eot.q1.c) & !is.na(eot.q3.c)) |  (!is.na(eot.q1.c) & is.na(eot.q3.c)), .N == 0],
  d[(is.na(b.q1.e) & !is.na(b.q3.e)) |  (!is.na(b.q1.e) & is.na(b.q3.e)), .N == 0],
  d[(is.na(b.q1.c) & !is.na(b.q3.c)) |  (!is.na(b.q1.c) & is.na(b.q3.c)), .N == 0],
  ## Check CI data:
  d[(is.na(eot.ci.low.e) & !is.na(eot.ci.high.e)) |  
      (!is.na(eot.ci.low.e) & is.na(eot.ci.high.e)), .N == 0],
  d[(is.na(eot.ci.low.c) & !is.na(eot.ci.high.c)) |  
      (!is.na(eot.ci.low.c) & is.na(eot.ci.high.c)), .N == 0],
  d[(is.na(b.ci.low.e) & !is.na(b.ci.high.e)) |  
      (!is.na(b.ci.low.e) & is.na(b.ci.high.e)), .N == 0],
  d[(is.na(b.ci.low.c) & !is.na(b.ci.high.c)) |  
      (!is.na(b.ci.low.c) & is.na(b.ci.high.c)), .N == 0],
  ## Check min-max data:
  d[(is.na(eot.min.e) & !is.na(eot.max.e)) |  (!is.na(eot.min.e) & is.na(eot.max.e)), .N == 0],
  d[(is.na(eot.min.c) & !is.na(eot.max.c)) |  (!is.na(eot.min.c) & is.na(eot.max.c)), .N == 0],
  d[(is.na(b.min.e) & !is.na(b.max.e)) |  (!is.na(b.min.e) & is.na(b.max.e)), .N == 0],
  d[(is.na(b.min.c) & !is.na(b.max.c)) |  (!is.na(b.min.c) & is.na(b.max.c)), .N == 0])
```

Check that $q_1 < median < q_3$ where they are non-missing. Perform similar checks for min, max, lower CI and upper CI values:
```{r}
stopifnot(
  ## Check Q1 and Q3 data:
  d[!is.na(eot.q1.e) & !is.na(eot.med.e), all(eot.q1.e <= eot.med.e & eot.med.e <= eot.q3.e)],
  d[!is.na(eot.q1.c) & !is.na(eot.med.c), all(eot.q1.c <= eot.med.c & eot.med.c <= eot.q3.c)],
  d[!is.na(b.q1.e) & !is.na(b.med.e), all(b.q1.e <= b.med.e & b.med.e <= b.q3.e)],
  d[!is.na(b.q1.c) & !is.na(b.med.c), all(b.q1.c <= b.med.c & b.med.c <= b.q3.c)],
  ## Check min-max data:
  d[!is.na(eot.q1.e) & !is.na(eot.min.e), all(eot.min.e < eot.q1.e & eot.q3.e < eot.max.e)],
  d[!is.na(eot.q1.c) & !is.na(eot.min.c), all(eot.min.c < eot.q1.c & eot.q3.c < eot.max.c)],
  d[!is.na(b.q1.e) & !is.na(b.min.e), all(b.min.e < b.q1.e & b.q3.e < b.max.e)],
  d[!is.na(b.q1.c) & !is.na(b.min.c), all(b.min.c < b.q1.c & b.q3.c < b.max.c)],
  ## Check CI data:
  d[!is.na(eot.ci.low.e) & !is.na(eot.m.e), all(eot.ci.low.e < eot.m.e & eot.m.e < eot.ci.high.e)],
  d[!is.na(eot.ci.low.c) & !is.na(eot.m.c), all(eot.ci.low.c < eot.m.c & eot.m.c < eot.ci.high.c)],
  d[!is.na(b.ci.low.e) & !is.na(b.m.e), all(b.ci.low.e < b.m.e & b.m.e < b.ci.high.e)],
  d[!is.na(b.ci.low.c) & !is.na(b.m.c), all(b.ci.low.c < b.m.c & b.m.c < b.ci.high.c)])
```


## Format data

Make copy of data with formatting applied:
```{r}
d2 <- copy(d)
```

### Compute SEM from CI

```{r}
## End-Of-Trial:
##  Experimental group:
d2[is.na(eot.sd.e) & is.na(eot.se.e) & !is.na(eot.ci.low.e) & !is.na(eot.ci.high.e), 
   eot.se.e := ci2se(ci=c(eot.ci.low.e, eot.ci.high.e), n = n.e), 
   by=row_id]
## Control group:
d2[is.na(eot.sd.c) & is.na(eot.se.c) & !is.na(eot.ci.low.c) & !is.na(eot.ci.high.c), 
   eot.se.c := ci2se(ci=c(eot.ci.low.c, eot.ci.high.c), n = n.c), 
   by=row_id]

## Baseline:
##  Experimental group:
d2[is.na(b.sd.e) & is.na(b.se.e) & !is.na(b.ci.low.e) & !is.na(b.ci.high.e), 
   b.se.e := ci2se(ci=c(b.ci.low.e, b.ci.high.e), n = n.e), 
   by=row_id]
## Control group:
d2[is.na(b.sd.c) & is.na(b.se.c) & !is.na(b.ci.low.c) & !is.na(b.ci.high.c), 
   b.se.c := ci2se(ci=c(b.ci.low.c, b.ci.high.c), n = n.c), 
   by=row_id]
```



### Compute SD from SEM

The relation between standard error (SE), standard deviation (SD) and sample size (n) is given by:
$$SE = SD / \sqrt{n}$$


```{r}
## End-Of-Trial:
##  Experimental group:
d2[is.na(eot.sd.e) & !is.na(eot.se.e), 
  eot.sd.e := eot.se.e * sqrt(n.e)]
##  Control group:
d2[is.na(eot.sd.c) & !is.na(eot.se.c), 
  eot.sd.c := eot.se.c * sqrt(n.c)]

## Baseline:
##  Experimental group:
d2[is.na(b.sd.e) & !is.na(b.se.e), 
  b.sd.e := b.se.e * sqrt(n.e)]
##  Control group:
d2[is.na(b.sd.c) & !is.na(b.se.c), 
  b.sd.c := b.se.c * sqrt(n.c)]

# Remove SE variables after converting available SE values to SD:
d2[, `:=`(eot.se.e = NULL,
          eot.se.c = NULL, 
          b.se.e   = NULL, 
          b.se.c   = NULL)]
```


### Impute mean and SD where missing from percentile data

Copy mean to median when sd is absent and only mean is present. This is to estimate SD using the median + range method in this case (approximating median with mean):
```{r}
d2[is.na(eot.sd.e) & is.na(eot.med.e) & !is.na(eot.m.e), eot.med.e := eot.m.e]
d2[is.na(eot.sd.c) & is.na(eot.med.c) & !is.na(eot.m.c), eot.med.c := eot.m.c]
d2[is.na(b.sd.e) & is.na(b.med.e) & !is.na(b.m.e), b.med.e := b.m.e]
d2[is.na(b.sd.c) & is.na(b.med.c) & !is.na(b.m.c), b.med.c := b.m.c]
```

Do the actual imputation of mean+SD for EOT and baseline separately using the `metacont` function from package **meta**:
```{r, warning=FALSE}
d2[, do_impute := fifelse(is.na(eot.m.e) | is.na(eot.sd.e), 1.0, 0.0)]

# Impute mean and SD where missing for EOT:
meta01 <- metacont(
  # Experimental group:
  n.e      = n.e, 
  mean.e   = eot.m.e, 
  sd.e     = eot.sd.e, 
  median.e = eot.med.e, 
  q1.e     = eot.q1.e, 
  q3.e     = eot.q3.e,
  min.e    = eot.min.e, 
  max.e    = eot.max.e,
  # Control group:
  n.c      = n.c, 
  mean.c   = eot.m.c, 
  sd.c     = eot.sd.c, 
  median.c = eot.med.c, 
  q1.c     = eot.q1.c, 
  q3.c     = eot.q3.c,
  min.c    = eot.min.c, 
  max.c    = eot.max.c,
  # Others arguments:
  studlab  = row_id,
  data     = d2,  
  sm       = "SMD") 

## Extract data on EOT:
d2b <- with(meta01, data.table(studlab, n.e, mean.e, sd.e, 
                               n.c, mean.c, sd.c))
# Rename variables to align with previous naming standard:
setnames(d2b, function(z) fifelse(grepl("^mean|^sd", z), paste0("eot.", z), z))
setnames(d2b, function(z) gsub("mean", "m", z, fixed = TRUE))

# Impute mean and SD where missing for Baseline:
meta01 <- metacont(
  # Experimental group:
  n.e      = n.e, 
  mean.e   = b.m.e, 
  sd.e     = b.sd.e, 
  median.e = b.med.e, 
  q1.e     = b.q1.e, 
  q3.e     = b.q3.e,
  min.e    = b.min.e, 
  max.e    = b.max.e,
  # Control group:
  n.c      = n.c, 
  mean.c   = b.m.c, 
  sd.c     = b.sd.c, 
  median.c = b.med.c, 
  q1.c     = b.q1.c, 
  q3.c     = b.q3.c,
  min.c    = b.min.c, 
  max.c    = b.max.c,
  # Others arguments:
  studlab  = row_id,
  data     = d2,
  sm       = "SMD") 

## Extract data on Baseline:
d2c <- with(meta01, data.table(studlab, n.e, mean.e, sd.e, 
                               n.c, mean.c, sd.c))
# Rename variables to align with previous naming standard:
setnames(d2c, function(z) fifelse(grepl("^mean|^sd", z), paste0("b.", z), z))
setnames(d2c, function(z) gsub("mean", "m", z, fixed = TRUE))
```

Merge imputed data into a completed dataset:
```{r}
## Merge data on EOT and Baseline:
stopifnot(nrow(d2) == nrow(d2b),
          nrow(d2) == nrow(d2c))
d2d <- merge(d2b, d2c[, -c("n.c", "n.e")], by="studlab", sort = FALSE)
stopifnot(length(grep("\\.x$|\\.y$", names(d2d))) == 0)

## Merge non-numeric columns with data-columns:
nm2 <- names(d2)
data_cols <- nm2[grep("\\.e$|\\.c$", nm2)]
d3 <- merge(d2[, -data_cols, with=FALSE], d2d,
            by.x="row_id", by.y="studlab", sort = FALSE)
stopifnot(length(grep("\\.x$|\\.y$", names(d3))) == 0,
          nrow(d2) == nrow(d3))
```

## Save data to disk

Datasets:

- `d`:  Validated source data before conversion of SE or CI to SD.
- `d2`:  Data after deriving missing SE values from confidence intervals and converting available SE values to SD.
- `d3`: Final analytic dataset after imputation of missing means and SDs from percentile data where possible.


Save validated and processed datasets:
```{r}
saveRDS(d, "data_derived/pico2_1.rds")
saveRDS(d2, "data_derived/pico2_2.rds")
saveRDS(d3, "data_derived/pico2_3.rds")
```


## Appendix

**R** session info:
```{r}
sessionInfo()
```

