---
title: "PICO 2: Overview Forest Plots for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d <- readRDS("data_derived/pico2_1.rds")
d3 <- readRDS("data_derived/pico2_3.rds")
```

```{r plot_function}
mk_plot <- function(meta_obj) {
  forest(
    meta_obj,
    digits.sd = 2,
    digits.mean = 2,
    xlim = c(-2, 2),
    print.subgroup.name = FALSE,
    overall = FALSE,
    overall.hetstat = FALSE,
    common = FALSE,
    random = FALSE,
    resid.hetstat = FALSE,
    test.overall.common = FALSE,
    test.overall.random = FALSE,
    test.subgroup = FALSE,
    test.effect.subgroup.fixed = FALSE,
    test.effect.subgroup.random = FALSE
  )

  invisible(NULL)
}
```


## Validated source data - EOT

```{r, results='hide', warning=FALSE}
meta01 <- metacont(n.e    = n.e, 
                   mean.e = eot.m.e, 
                   sd.e   = eot.sd.e, 
                   n.c    = n.c, 
                   mean.c = eot.m.c, 
                   sd.c   = eot.sd.c, 
                   studlab=paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data=d,
                   sm="SMD") 
```

Display validated source data before conversion of SE or CI to SD and before imputation of mean or SD:

```{r ow_forestplot01_eot, fig.width=12, fig.height=35}
mk_plot(meta01)
```


## Display imputed data - EOT

Display the final analytic data after conversion of available SE and CI information to SD and imputation of missing means and SDs where possible:

```{r, warning=FALSE}
meta01 <- metacont(n.e    = n.e, 
                   mean.e = eot.m.e, 
                   sd.e   = eot.sd.e, 
                   n.c    = n.c, 
                   mean.c = eot.m.c, 
                   sd.c   = eot.sd.c, 
                   studlab=paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data=d3,
                   sm="SMD") 
```

```{r ow_forestplot02_eot, fig.width=12, fig.height=40}
mk_plot(meta01)
```


## Validated source data - Baseline

```{r, results='hide', warning=FALSE}
meta01 <- metacont(n.e    = n.e, 
                   mean.e = b.m.e, 
                   sd.e   = b.sd.e, 
                   n.c    = n.c, 
                   mean.c = b.m.c, 
                   sd.c   = b.sd.c, 
                   studlab=paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data=d,
                   sm="SMD") 
```

Display all original data after conversion of SEM to SD but before imputation of mean or SD:

```{r ow_forestplot01_b, fig.width=12, fig.height=33}
mk_plot(meta01)
```


## Display imputed data - Baseline

Display data after imputation of mean and SD:

```{r, warning=FALSE}
meta01 <- metacont(n.e    = n.e, 
                   mean.e = b.m.e, 
                   sd.e   = b.sd.e, 
                   n.c    = n.c, 
                   mean.c = b.m.c, 
                   sd.c   = b.sd.c, 
                   studlab=paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data=d3,
                   sm="SMD") 
```

```{r ow_forestplot02_b, fig.width=12, fig.height=40}
mk_plot(meta01)
```


## Appendix

**R** session info:
```{r}
sessionInfo()
```







