---
title: "PICO 2: Leave One Out Influence Analyses"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.retina = 4, dev = c('png', 'pdf'))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Leave-one-out analyses - Psychometrics

### Read data

```{r}
d3 <- readRDS("data_derived/pico2_3.rds")

d4 <- d3[outcome.cat == "psychometrics" & include == 1]
d4[, grp := .GRP, by=.(author, year, outcome, disease)]
d4 <- d4[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d4, -year)
```


Outcome sub-categories with more than 1 study:
```{r}
tab <- d4[, .N, outcome.subcat][N > 1][order(-N)]
uni_cats <- tab[, outcome.subcat] 
tab 
```

### Define function

```{r function_mk_leave1out}
mk_leave1out <- function(data, width=12, height=10, res=100, 
                         save_fig = interactive(), fig_path="figures/test01.png") {
  pdata <- copy(data)

  meta01 <- metacont(n.e     = n.e, 
                     mean.e  = eot.m.e * direction, 
                     sd.e    = eot.sd.e, 
                     n.c     = n.c, 
                     mean.c  = eot.m.c * direction, 
                     sd.c    = eot.sd.c, 
                     studlab = paste0(author, " ", year), 
                     data    = pdata, 
                     sm      = "SMD")
  
  if(save_fig) png(fig_path, width=width, height=height, unit="in", res=res)
  forest(metainf(meta01, pooled = "random"))
  if(save_fig) dev.off()
}
```


### `r uni_cats[1]`

```{r leave1out_1, fig.width=9, fig.height=3}
plotdata <- d4[outcome.subcat == uni_cats[1]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[2]`

```{r leave1out_2, fig.width=9, fig.height=3}
plotdata <- d4[outcome.subcat == uni_cats[2]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[3]`

```{r leave1out_3, fig.width=9, fig.height=3}
plotdata <- d4[outcome.subcat == uni_cats[3]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[4]`

```{r leave1out_4, fig.width=9, fig.height=3}
plotdata <- d4[outcome.subcat == uni_cats[4]]
mk_leave1out(plotdata, width = 9, height = 3)
```


## Leave-one-out analyses - Biomarkers

### Prepare data

```{r}
d4 <- d3[
  outcome.cat != "psychometrics" &
  include == 1 &
  !is.na(eot.m.e) &
  !is.na(eot.sd.c) &
  eot.sd.c > 0
]

d4[, grp := .GRP, by = .(author, year, outcome, disease)]
d4 <- d4[, .SD[which.min(abs(duration - followup))], by = grp]
setorder(d4, -year)

d4[, n_studies := .N, keyby=outcome]
d5 <- d4[n_studies > 1] |> droplevels()
outcome_levels <- d5[, .N, outcome][order(-N)][, outcome]
d5[, f_outcome := factor(outcome, levels = outcome_levels)]
d5[, .N, keyby = f_outcome]
```


Biomarker outcomes with more than 1 study:
```{r}
tab <- d5[, .N, f_outcome][N > 1][order(-N)]
uni_cats <- tab[, f_outcome] 
tab 
```

### `r uni_cats[1]`

```{r leave1out_biom1, fig.width=9, fig.height=5}
plotdata <- d5[outcome == uni_cats[1]]
mk_leave1out(plotdata, width = 9, height = 5)
```

### `r uni_cats[2]`

```{r leave1out_biom2, fig.width=9, fig.height=5}
plotdata <- d5[outcome == uni_cats[2]]
mk_leave1out(plotdata, width = 9, height = 5)
```

### `r uni_cats[3]`

```{r leave1out_biom3, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[3]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[4]`

```{r leave1out_biom4, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[4]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[5]`

```{r leave1out_biom5, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[5]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[6]`

```{r leave1out_biom6, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[6]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[7]`

```{r leave1out_biom7, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[7]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[8]`

```{r leave1out_biom8, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[8]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[9]`

```{r leave1out_biom9, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[9]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[10]`

```{r leave1out_biom10, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[10]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[11]`

```{r leave1out_biom11, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[11]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[12]`

```{r leave1out_biom12, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[12]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[13]`

```{r leave1out_biom13, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[13]]
mk_leave1out(plotdata, width = 9, height = 3)
```

### `r uni_cats[14]`

```{r leave1out_biom14, fig.width=9, fig.height=3}
plotdata <- d5[outcome == uni_cats[14]]
mk_leave1out(plotdata, width = 9, height = 3)
```
