---
title: "PICO 3: Prepare Data for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

## Create folders if missing:
root_dir <- knitr::opts_knit$get("root.dir")
figs_dir <- paste0(root_dir, "/figures")
results_dir <- paste0(root_dir, "/results")
data_derived_dir <- paste0(root_dir, "/data_derived")
if(!dir.exists(figs_dir))  dir.create(figs_dir)
if(!dir.exists(results_dir)) dir.create(results_dir)
if(!dir.exists(data_derived_dir)) dir.create(data_derived_dir)
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(readxl)
library(data.table)
library(writexl)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

Check existence of file with raw data:
```{r}
RAWDATA <- "data_raw/raw_data_butyrate_meta_analysis.xlsx"
stopifnot(file.exists(RAWDATA))
```

Preliminary reading of data treating only empty cells as `NA` (cells containing "NA" are _not_ treated as `NA`):
```{r, results='hide'}
## Reading data doesn't work of the file is open in desktop application. Opening it in the 
## browser application works fine, though.
d0 <- quick_read_excel(RAWDATA, na = c(""), sheet="PICO 3", guess_max = 1e4)
setDT(d0)
## Preview dataset (only shown in interactive sessions):
dim(d0)
str(d0)
head(d0)
tail(d0)
```


### Check for expected columns (=variables) in dataset

```{r}
expected_col_names <- 
  c("include", "outcome.cat", "outcome.sub.cat", "outcome", "direction", 
    "o.units", "author", "year", "model.cat", "model.spec", "intervention.type", 
    "dosis", "i.units", "duration", "followup", "control", "n.e", "m.e","sd.e", 
    "sem.e", "med.e", "q1.e", "q3.e", "combine", "n.c", "m.c", "sd.c", "sem.c", 
    "med.c", "q1.c", "q3.c","analysis.designation", "extraction.source", "notes")

if(!all(expected_col_names %in% names(d0))) {
  missing_var <- setdiff(expected_col_names, names(d0))
  stop("Required variables in data not available: ", paste(missing_var, collapse = ", "))
}
if(length(new_variables <- setdiff(names(d0), expected_col_names))) 
  warning("New variables detected in data: ", paste(new_variables, collapse = ", "))
```


### Check for empty cells in dataset 

We require no missing cells in the dataset.

```{r}
## Only read read columns until free-text columns (notes/comments):
max_col <- min(grep("notes|Notes|comment|Comment", names(d0))) - 1
d <- d0[, 1:max_col]

# Columns in which to ignore missing cells:
col_ignore <- character(0L)
stopifnot(all(col_ignore %in% names(d)))

## Identify cells with missing values:
x <- sapply(d[, -col_ignore, with = FALSE], function(x) sum(is.na(x)))
x <- x[x > 0]

if(length(x) > 0) {
  # Get list of places with empty cells:
  nm <- names(x)
  message("Missing cells in the following columns (row numbers):")
  print(lapply(setNames(nm, nm), function(n) which(is.na(d0[, get(n)]))+1) )
  warning("Missing cells are not allowed!")
}
```

### Check for "return/newline" line endings in data

```{r}
# Find which columns contain \r (and/or \r\n):
cols <- d[, lapply(.SD, function(x) any(grepl("\r", x, fixed = TRUE))), .SDcols = is.character] |>
  unlist() |> which() |> names()
if(length(cols) > 0) {
  message("Cells containing 'return/newline' in the following columns (row numbers):")
  print(lapply(d[, .SD, .SDcols = cols], function(x) 
    which(grepl("\r", x, fixed = TRUE)) + 1))
  stop("Fix data before proceeding!")
}
```

## Read and check data

Read the data treating `NA` as missing:
```{r, results='hide'}
# Read file again including only real data-rows:
d0 <- quick_read_excel(RAWDATA, na = c("", "NA"), guess_max = 1e4,
                       sheet="PICO 3") 
setDT(d0)
d <- d0[, 1:max_col][, -col_ignore, with = FALSE]
dim(d)
head(d)
tail(d)
str(d, list.len=1e3)
```

Prepend a row id column:
```{r}
d[, row_id := 1:.N]
setcolorder(d, "row_id")
```


### Check that all data columns are numeric 

Data columns end in `.e` or `.c`:
```{r}
nm <- names(d)
data_cols <- nm[grep("\\.e$|\\.c$", nm)]
not_is_numeric <- names(which(!sapply(d[, ..data_cols], is.numeric)))
## Coerce non-numeric data-variables to numeric:
for(nam in not_is_numeric) { # nam <- not_is_numeric[1]
  tmp <- d[, get(nam)]
  tmp_num <- suppressWarnings(as.numeric(tmp))
  if(sum(is.na(tmp_num)) > sum(is.na(tmp))) 
    warning("NAs introduced when coercing '", nam, "' to numeric")
  d[, (nam) := as.numeric(get(nam))]
}
```


### Tabulate non-numeric variables 

```{r}
nms <- nm[!sapply(d, is.numeric)]
tables <- lapply(setNames(nms, nms), function(col) {
  totpct(d[, .N, by=col][order(-N)])
})
write_xlsx(tables, path="results/PICO3_categorical_variable_tables.xlsx")
```


### Check consistency in reporting of mean/median/sd/se

Check that if mean/median is reported in the experimental group, it is also reported in the control group. This only applies to non-combine groups.
```{r}
tmp <- d[is.na(combine)]
stopifnot(tmp[is.na(m.e) & !is.na(m.c) | !is.na(m.e) & is.na(m.c), .N == 0], # *
  tmp[is.na(sd.e) & !is.na(sd.c) | !is.na(sd.e) & is.na(sd.c), .N == 0],
  tmp[is.na(sem.e) & !is.na(sem.c) | !is.na(sem.e) & is.na(sem.c), .N == 0] # *
)
```


### Check consistency of quartile variables
Check that `q1` and `q3` are either both reported or both missing, and that reported medians lie between the first and third quartiles. 
```{r} 
stopifnot(
  ## Check Q1 and Q3 data:
  d[(is.na(q1.e) & !is.na(q3.e)) |  (!is.na(q1.e) & is.na(q3.e)), .N == 0],
  d[(is.na(q1.c) & !is.na(q3.c)) |  (!is.na(q1.c) & is.na(q3.c)), .N == 0])
```

Check that $q_1 < \mathrm{median} < q_3$ where they are non-missing. Perform similar checks for min, max, lower CI and upper CI values:
```{r} 
stopifnot(
  ## Check Q1 and Q3 data:
  d[!is.na(q1.e) & !is.na(med.e), all(q1.e <= med.e & med.e <= q3.e)],
  d[!is.na(q1.c) & !is.na(med.c), all(q1.c <= med.c & med.c <= q3.c)])
```


## Format data

Make copy of data with formatting applied:
```{r}
d2 <- copy(d)
```


### Compute SD from SEM

The relation between standard error (SE), standard deviation (SD) and sample size (n) is given by:
$$SE = SD / \sqrt{n}$$


```{r}
## End-Of-Trial:
##  Experimental group:
d2[is.na(sd.e) & !is.na(sem.e), 
  sd.e := sem.e * sqrt(n.e)]
##  Control group:
d2[is.na(sd.c) & !is.na(sem.c), 
  sd.c := sem.c * sqrt(n.c)]

# Remove 'sem'-variables:
d2[, `:=`(sem.e = NULL,
          sem.c = NULL)]
```


### Impute mean and SD where missing from percentile data

Copy mean to median when sd is absent and only mean is present. This is to estimate SD using the median + range method in this case (approximating median with mean):
```{r} 
d2[is.na(sd.e) & is.na(med.e) & !is.na(m.e), med.e := m.e]
d2[is.na(sd.c) & is.na(med.c) & !is.na(m.c), med.c := m.c]
```

Do the actual imputation of mean+SD using the `metacont` function from package **meta**:
```{r, warning=FALSE}
d2[, do_impute := fifelse(is.na(m.e) | is.na(sd.e), 1.0, 0.0)]

# Impute missing means and SDs:
meta01 <- metacont(
  # Experimental group:
  n.e      = n.e, 
  mean.e   = m.e, 
  sd.e     = sd.e, 
  median.e = med.e, 
  q1.e     = q1.e, 
  q3.e     = q3.e,

  # Control group:
  n.c      = n.c, 
  mean.c   = m.c, 
  sd.c     = sd.c, 
  median.c = med.c, 
  q1.c     = q1.c, 
  q3.c     = q3.c,

  # Others arguments:
  studlab  = row_id,
  data     = d2, 
  sm       = "SMD") 

## Extract data:
d2b <- with(meta01, data.table(studlab, n.e, mean.e, sd.e,
                          n.c, mean.c, sd.c))
setnames(d2b, function(z) gsub("mean", "m", z, fixed = TRUE))
```

Merge imputed data into a completed dataset:
```{r}
## Merge non-numeric columns with data-columns:
nm2 <- names(d2)
data_cols <- nm2[grep("\\.e$|\\.c$", nm2)]
d3 <- merge(d2[, -data_cols, with=FALSE], d2b,
            by.x="row_id", by.y="studlab", sort = FALSE)
stopifnot(length(grep("\\.x$|\\.y$", names(d3))) == 0,
          nrow(d2) == nrow(d3))
```


## Collapse control groups

First identify rows of data that needs combining, then loop over unique `combine` values and separately combine values in experimental and control groups. Lastly, merge combine data with data rows that did not need combining to achieve a final complete dataset:
```{r}
## Split data into section that do not and that do needs combining:
d3a <- d3[is.na(combine)]
d3b <- d3[!is.na(combine)]

## Collapse relevant values:
combine_values <- d3b[, unique(combine)]
d3c <- lapply(combine_values, function(cv) { 
  dd <- d3b[combine == cv]
  res <- dd[!is.na(n.c) & !is.na(n.e)][1]
  # Each combine identifier must produce exactly one analytic comparison:
  stopifnot(nrow(res) == 1)
  
  ## Collapse experimental groups:
  dde <- dd[!is.na(n.e)]
  missings <- sum(!complete.cases(dde[, .(n.e, m.e, sd.e)]))
  if(nrow(dde) > 1 & missings == 0) {
   res.e <- dde[, collapse_groups(ns = n.e, m.e, sd.e)]
    res[, `:=`(n.e      = res.e["n"],
               m.e  = res.e["mu"],
               sd.e = res.e["sd"])]
  }
  
  ## Collapse control groups for multi-arm studies:
  ddc <- dd[!is.na(n.c)]
  missings <- sum(!complete.cases(ddc[, .(n.c, m.c, sd.c)]))
  if(nrow(ddc) > 1 & missings == 0) {
    res.c <- ddc[, collapse_groups(ns = n.c, m.c, sd.c)]
    res[, `:=`(n.c      = res.c["n"],
               m.c  = res.c["mu"],
               sd.c = res.c["sd"])]
  }
  res
}) |> rbindlist()
## Verify that each combine identifier produced one analytic comparison: 
stopifnot(nrow(d3c) == length(combine_values))

## Merge section that did not need combining with combine section:
d4 <- rbind(d3a, d3c)[order(row_id)]
```


## Save data to disk

Datasets:

- `d`:  Original dataset (validated)
- `d2`: Data with SD imputed from SEM (and CI)
- `d3`: Data with mean and SD imputed from percentile data where otherwise unavailable
- `d4`: Data where data sections are combined

Preview `d4`. Note that we only keep N, mean, and SD (for EOT & baseline, and for experimental and control groups):
```{r}
d4
```

Save data files:
```{r}
saveRDS(d, "data_derived/pico3_1.rds")
saveRDS(d2, "data_derived/pico3_2.rds")
saveRDS(d3, "data_derived/pico3_3.rds")
saveRDS(d4, "data_derived/pico3_4.rds")
```


## Appendix

**R** session info:
```{r}
sessionInfo()
```

