---
title: "PICO 3: Overview Forest Plots for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d <- readRDS("data_derived/pico3_1.rds")
d3 <- readRDS("data_derived/pico3_3.rds")
d4 <- readRDS("data_derived/pico3_4.rds")
```

```{r plot_function}
mk_plot <- function(meta_obj, width=12, height=10, res=100, 
                    save_fig = interactive(), fig_path="figures/test01.png", ...) {
  if(save_fig) png(fig_path, width=width, height=height, unit="in", res=res)
  forest(meta_obj, 
         digits.sd = 2, 
         digits.mean = 2, 
         xlim=c(-2, 2),
         print.subgroup.name = FALSE,
         overall = FALSE, 
         overall.hetstat = FALSE,
         common = FALSE,
         random = FALSE,
         resid.hetstat = FALSE,
         test.overall.common = FALSE,
         test.overall.random = FALSE,
         test.subgroup = FALSE,
         test.effect.subgroup.fixed = FALSE, 
         test.effect.subgroup.random = FALSE)
         
  if(save_fig) dev.off()
  invisible(NULL)
}
```


## Display original data 

```{r, results='hide', warning=FALSE}
meta01 <- metacont(n.e       = n.e, 
                   mean.e    = m.e, 
                   sd.e      = sd.e, 
                   n.c       = n.c, 
                   mean.c    = m.c, 
                   sd.c      = sd.c, 
                   studlab   = paste0(author, " ", year, ", ", outcome), 
                   subgroup  = outcome.cat,
                   data      = d,
                   sm        = "SMD") 
```

Display all original data:

```{r ow_forestplot01_eot, fig.width=15, fig.height=107}
mk_plot(meta01, width=15, height=107)
```


## Display imputed data 

Display data after conversion of SEM to SD and imputation of missing means and SDs, but before combination of intervention arms in multi-arm studies:

```{r, warning=FALSE}
meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data     = d3,
                   sm       = "SMD") 
```

```{r ow_forestplot02_eot, fig.width=15, fig.height=107}
mk_plot(meta01, width=15, height=107)
```


## Display imputed and combined data / Final analytic data

Display the final analytic data after conversion of SEM to SD, imputation of missing means and SDs, and combination of intervention arms in multi-arm studies. Each combined comparison contains complete intervention and control data, allowing all included outcomes to be represented in the forest plot:

```{r, warning=FALSE}
meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year, ", ", outcome), 
                   subgroup = outcome.cat,
                   data     = d4,
                   sm       = "SMD") 
```

```{r ow_forestplot03_eot, fig.width=15, fig.height=90}
mk_plot(meta01, width=15, height=90)
```



## Appendix

**R** session info:
```{r}
sessionInfo()
```







