---
title: "PICO 3: Main Forest Plots for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d4 <- readRDS("data_derived/pico3_4.rds")
```

```{r functions}
rescale_means <- function(object) {
  object$mean.e <- abs(object$mean.e)
  object$mean.c <- abs(object$mean.c)
  object
}
mk_plot <- function(meta_obj, width=12, height=10, res=100, 
                    save_fig = interactive(), fig_path="figures/test01.png", ...) {
  if(save_fig) png(fig_path, width=width, height=height, unit="in", res=res)
  
  forest(rescale_means(meta_obj), 
         digits.sd = 1, 
         digits.mean = 2, 
         xlim=c(-2, 2), 
         leftcols = c("studlab", "outcome", "model.cat", "n.e", "mean.e", "sd.e", 
                      "n.c", "mean.c", "sd.c"),
         leftlabs = c("Study", "Outcome measure", "Disease model", "Total", 
                      "Mean", "SD", "Total", "Mean", "SD"),
         just.addcols.left = "right",
         rightcols = c("effect", "ci", "w.random"),
         
         # Overall/Random/Common effect settings:
         overall = FALSE,
         common = FALSE,
         random = TRUE,
         overall.hetstat = FALSE,
         hetstat = TRUE,
         test.overall.random = FALSE,
         
         # Subgroup settings:
         subgroup = TRUE,
         print.subgroup.name = FALSE,
         random.subgroup = TRUE,
         test.subgroup = FALSE,
         test.effect.subgroup = TRUE,
         label.test.effect.subgroup.random = "Test of random effect: ",
         
         # Prediction interval settings:
         prediction = FALSE,
         prediction.subgroup = TRUE,
         ...

         )
  
  if(save_fig) dev.off()
  invisible(NULL)
}
```

# All data

## Behavioral data

Extract data for the behavioral outcome category. For each combination of study, outcome and disease, select the data row where follow-up is closest to duration of the trial:
```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)

# Add * to study-label for duplicated studies:
d5[, studlab := paste0(author, " ", year), by = .I]
d5[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d5[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=50}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=50)
```


## Behavioral data - |SMD| < 2

```{r}
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]

# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=40}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d6,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=40)
```



## Behavioral data - |SMD| < 1.6

```{r}
d6 <- d5[abs(SMD) < 1.6]

# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=36}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d6,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=36)
```


# Excluding graph extracted data - primary behavioral analysis


## Behavioral data

Extract data for the behavioral outcome category. For each combination of study, outcome and disease, select the data row where follow-up is closest to duration of the trial:
```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1 & 
           extraction.source != "graph_extracted_data"]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)

# Add * to study-label for duplicated studies:
d5[, studlab := paste0(author, " ", year), by = .I]
d5[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d5[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=20}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=20)
```


## Behavioral data - |SMD| < 2

```{r}
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]

# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=19}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d6,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=19)
```



## Behavioral data - |SMD| < 1.6

```{r}
d6 <- d5[abs(SMD) < 1.6]

# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]
```


```{r fig.width=16, fig.height=16}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d6,
                     sm       = "SMD") 
mk_plot(meta_obj, colgap.forest.left = "0.5cm",
       colgap.forest.right = "0.5cm",
       fontsize = 11, width=16, height=16)
```


## Appendix

**R** session info:
```{r}
sessionInfo()
```












