---
title: "PICO 3: Funnel Plots"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.retina = 4, dev = c('png', 'pdf'))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d4 <- readRDS("data_derived/pico3_4.rds")

d5 <- d4[outcome.cat == "behavioral" & include == 1]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)
```


# All included data

Outcome sub-categories with more than 1 study/one effect estimate (some preclinical studies have different mice cohorts for the same studies, marked with * at the forest plots):
```{r}
tab <- d5[, .N, outcome.sub.cat][N > 1][order(-N)]
uni_cats <- tab[, outcome.sub.cat] 
tab 
```
## Behavioral outcomes
### Funnel plots & Egger's test


```{r funnel_n_function}
mk_funnelplot <- function(data, placement = "topleft") {
  pdata <- copy(data)

  meta01 <- metacont(n.e     = n.e, 
                     mean.e  = m.e * direction, 
                     sd.e    = sd.e, 
                     n.c     = n.c, 
                     mean.c  = m.c * direction, 
                     sd.c    = sd.c, 
                     studlab = paste0(author, " ", year), 
                     data    = pdata, 
                     sm      = "SMD")
  
  cc <- funnel(meta01, random = TRUE, common = FALSE,
               level = 0.95, contour = c(0.9, 0.95, 0.99))$col.contour
  legend(placement,legend = c("0.1 > p > 0.05", "0.05 > p > 0.01", "< 0.01"), 
         fill = cc, bg = "white")
  if(meta01$k > 9) metabias(meta01, k.min = 9) |> print()
  invisible(meta01)
}
```

```{r, eval=FALSE}
print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", sep="")
    cat("```{r funnel_", i, ", fig.width=7, fig.height=7}\n", sep="")
    cat("plotdata <- d5[outcome.sub.cat == uni_cats[", i, "]]\n", sep="")
    cat("mk_funnelplot(plotdata)\n")
    cat("```\n")
  }
}

print_code(length(uni_cats))

```


### `r uni_cats[1]`

```{r funnel_1, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[1]]
mk_funnelplot(plotdata)
```

### `r uni_cats[2]`

```{r funnel_2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[2]]
mk_funnelplot(plotdata)
```

### `r uni_cats[3]`

```{r funnel_3, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[3]]
mk_funnelplot(plotdata)
```

### `r uni_cats[4]`

```{r funnel_4, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[4]]
mk_funnelplot(plotdata)
```


### `r uni_cats[5]`

```{r funnel_5, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[5]]
mk_funnelplot(plotdata)
```

### `r uni_cats[6]`

```{r funnel_6, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[6]]
mk_funnelplot(plotdata)
```

### `r uni_cats[7]`

```{r funnel_7, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[7]]
mk_funnelplot(plotdata)
```

## Biomarker outcomes
### Funnel plots & Egger's test

```{r}
d5 <- d4[
  outcome.cat == "biomarkers" & include == 1 &
  !is.na(m.e) &
  !is.na(sd.c) &
  sd.c > 0
]

d5[, grp := .GRP, by = .(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by = grp]
setorder(d5, -year)

d5[, n_studies := .N, keyby=outcome]
d6 <- d5[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```


Outcome sub-categories with more than 1 study/effect estimate:
```{r}
tab <- d5[, .N, outcome][N > 1][order(-N)]
uni_cats <- tab[, outcome] 
tab 
```

```{r, eval=FALSE}
print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", sep="")
    cat("```{r funnel_", i, "biomark , fig.width=7, fig.height=7}\n", sep="")
    cat("plotdata <- d5[outcome == uni_cats[", i, "]]\n", sep="")
    cat("mk_funnelplot(plotdata)\n")
    cat("```\n")
  }
}

print_code(length(uni_cats))

```


### `r uni_cats[1]`

```{r funnel_1biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[1]]
mk_funnelplot(plotdata)
```

### `r uni_cats[2]`

```{r funnel_2biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[2]]
mk_funnelplot(plotdata)
```

### `r uni_cats[3]`

```{r funnel_3biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[3]]
mk_funnelplot(plotdata)
```

### `r uni_cats[4]`

```{r funnel_4biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[4]]
mk_funnelplot(plotdata)
```

### `r uni_cats[5]`

```{r funnel_5biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[5]]
mk_funnelplot(plotdata)
```

### `r uni_cats[6]`

```{r funnel_6biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[6]]
mk_funnelplot(plotdata)
```

### `r uni_cats[7]`

```{r funnel_7biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[7]]
mk_funnelplot(plotdata)
```

### `r uni_cats[8]`

```{r funnel_8biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[8]]
mk_funnelplot(plotdata)
```

### `r uni_cats[9]`

```{r funnel_9biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[9]]
mk_funnelplot(plotdata)
```

### `r uni_cats[10]`

```{r funnel_10biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[10]]
mk_funnelplot(plotdata)
```

### `r uni_cats[11]`

```{r funnel_11biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[11]]
mk_funnelplot(plotdata)
```

### `r uni_cats[12]`

```{r funnel_12biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[12]]
mk_funnelplot(plotdata)
```

### `r uni_cats[13]`

```{r funnel_13biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[13]]
mk_funnelplot(plotdata)
```

### `r uni_cats[14]`

```{r funnel_14biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[14]]
mk_funnelplot(plotdata)
```

### `r uni_cats[15]`

```{r funnel_15biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[15]]
mk_funnelplot(plotdata)
```

### `r uni_cats[16]`

```{r funnel_16biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[16]]
mk_funnelplot(plotdata)
```

### `r uni_cats[17]`

```{r funnel_17biomark , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[17]]
mk_funnelplot(plotdata)
```


# Excluding graph extracted data

```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1 & 
               extraction.source != "graph_extracted_data"]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)
```

Outcome sub-categories with more than 1 study:
```{r}
tab <- d5[, .N, outcome.sub.cat][N > 1][order(-N)]
uni_cats <- tab[, outcome.sub.cat] 
tab 
```

## Behavioral outcomes
### Funnel plots & Egger's test


```{r, eval=FALSE}
print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", sep="")
    cat("```{r funnel_", i, ", fig.width=7, fig.height=7}\n", sep="")
    cat("plotdata <- d5[outcome.sub.cat == uni_cats[", i, "]]\n", sep="")
    cat("mk_funnelplot(plotdata)\n")
    cat("```\n")
  }
}

print_code(length(uni_cats))

```


### `r uni_cats[1]`

```{r funnel_1_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[1]]
mk_funnelplot(plotdata)
```

### `r uni_cats[2]`

```{r funnel_2_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[2]]
mk_funnelplot(plotdata)
```

### `r uni_cats[3]`

```{r funnel_3_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[3]]
mk_funnelplot(plotdata)
```

### `r uni_cats[4]`

```{r funnel_4_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[4]]
mk_funnelplot(plotdata)
```

### `r uni_cats[5]`

```{r funnel_5_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[5]]
mk_funnelplot(plotdata)
```

### `r uni_cats[6]`

```{r funnel_6_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[6]]
mk_funnelplot(plotdata)
```

### `r uni_cats[7]`

```{r funnel_7_v2, fig.width=7, fig.height=7}
plotdata <- d5[outcome.sub.cat == uni_cats[7]]
mk_funnelplot(plotdata)
```

## Biomarker outcomes
### Funnel plots & Egger's test

```{r}
d5 <- d4[
  outcome.cat == "biomarkers" & include == 1 &
    extraction.source != "graph_extracted_data" & 
  !is.na(m.e) &
  !is.na(sd.c) &
  sd.c > 0
]

d5[, grp := .GRP, by = .(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by = grp]
setorder(d5, -year)

d5[, n_studies := .N, keyby=outcome]
d6 <- d5[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```


Outcome sub-categories with more than 1 study:
```{r}
tab <- d5[, .N, outcome][N > 1][order(-N)]
uni_cats <- tab[, outcome] 
tab 
```

```{r, eval=FALSE}
print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", sep="")
    cat("```{r funnel_", i, "biomark , fig.width=7, fig.height=7}\n", sep="")
    cat("plotdata <- d5[outcome == uni_cats[", i, "]]\n", sep="")
    cat("mk_funnelplot(plotdata)\n")
    cat("```\n")
  }
}

print_code(length(uni_cats))

```


### `r uni_cats[1]`

```{r funnel_1biomark_v2 , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[1]]
mk_funnelplot(plotdata)
```

### `r uni_cats[2]`

```{r funnel_2biomark_v2 , fig.width=7, fig.height=7}
plotdata <- d5[outcome == uni_cats[2]]
mk_funnelplot(plotdata)
```




## Appendix

**R** session info:
```{r}
sessionInfo()
```
















