---
title: "PICO 3: Tables for Meta Analysis"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r setup, include=FALSE}
library(data.table)
library(meta)
library(grid)
library(ggplot2)
```

```{r, setup02, include=FALSE}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d4 <- readRDS("data_derived/pico3_4.rds")
```

# All data

## Behavioral

### Table of SMD for each behavioral construct ($k > 1$, ie., all studies)

Extract behavioral outcome data. For each combination of study, outcome, disease model and intervention type, select the row where follow-up is closest to the intervention duration.

```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)

# Add * to study-label for duplicated studies:
d5[, studlab := paste0(author, " ", year), by = .I]
d5[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d5[, studlab_star := fifelse(studlab_n > 1, "*", "")]

d5[, n_studies := .N, keyby=outcome.sub.cat]
d6 <- d5[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome.sub.cat,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)

```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome.sub.cat  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome.sub.cat]
X <- as.data.table(merge(X1, X2, by="outcome.sub.cat"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_ps, fig.width=9, fig.height=3}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome.sub.cat, data = X) 

if(interactive())
  png("figures/test02.png", width=9, height=3, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


## Behavioral - |SMD| < 2

### Table of SMD for each behavioral constructs ($k > 1$)



```{r}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]
d6[, n_studies := .N, keyby=outcome.sub.cat]
d6 <- d6[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome.sub.cat,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome.sub.cat  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome.sub.cat]
X <- as.data.table(merge(X1, X2, by="outcome.sub.cat"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_ps_smd2, fig.width=9, fig.height=3}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome.sub.cat, data = X) 
if(interactive())
  png("figures/test02.png", width=9, height=3, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


## Biomarkers

### Table of SMD for each  biomarker outcome ($k > 1$)

Extract biomarker outcome data. For each combination of study, outcome, disease model and intervention type, select the row where follow-up is closest to the intervention duration:

```{r}
d5 <- d4[
  outcome.cat == "biomarkers" & include == 1 &
  !is.na(m.e) &
  !is.na(sd.c) &
  sd.c > 0
]

d5[, grp := .GRP, by = .(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by = grp]
setorder(d5, -year)

d5[, n_studies := .N, keyby=outcome]
d6 <- d5[n_studies > 1]


meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_bio, fig.width=9, fig.height=5}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X) 

if(interactive())
  png("figures/test02.png", width=9, height=5, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("Outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


### Table of SMD for each  biomarker outcome ($k = 1$)

Select biomarker outcomes represented by a single study or effect estimate ($k = 1$):

```{r}
d6 <- d5[n_studies == 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1b_bio, fig.width=10, fig.height=4}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X)

if(interactive())
  png("figures/test02.png", width=10, height=4, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("Outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


### SMD table with all studies (any $k$)

```{r}
meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d5, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d5[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
X <- X[Studies > 0]
```

### Forest plot of table 

```{r table_1_suppl_bio, fig.width=10, fig.height=7}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X)

if(interactive())
  png("figures/test02.png", width=10, height=7, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


# Excluding graph extracted data

## Behavioral

### SMD table for non-digitized behavioral constructs ($k > 1$)

Extract behavioral non-digitized (cf. extraction.source) outcome data. For each combination of study, outcome, disease model and intervention type, select the row where follow-up is closest to the intervention duration.

```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1 & 
               extraction.source != "graph_extracted_data"]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
setorder(d5, -year)

# Add * to study-label for duplicated studies:
d5[, studlab := paste0(author, " ", year), by = .I]
d5[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d5[, studlab_star := fifelse(studlab_n > 1, "*", "")]

d5[, n_studies := .N, keyby=outcome.sub.cat]
d6 <- d5[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome.sub.cat,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome.sub.cat  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome.sub.cat]
X <- as.data.table(merge(X1, X2, by="outcome.sub.cat"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_ps_v2, fig.width=9, fig.height=3}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome.sub.cat, data = X)

if(interactive())
  png("figures/test02.png", width=9, height=3, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


## Behavioral - |SMD| < 2

### Table of SMD for behavioral non-digitized construct ($k > 1$)

Extract behavioral non-digitized outcome data with |SMD| < 2:

```{r}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = paste0(studlab, studlab_star),
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]
d6[, n_studies := .N, keyby=outcome.sub.cat]
d6 <- d6[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome.sub.cat,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome.sub.cat  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
# X1
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome.sub.cat]
X <- as.data.table(merge(X1, X2, by="outcome.sub.cat"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_ps_smd2_v2, fig.width=9, fig.height=3}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome.sub.cat, data = X) 

if(interactive())
  png("figures/test02.png", width=9, height=3, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


## Biomarkers

### Table of SMD for each  biomarker outcome ($k > 1$)

Extract data for all non-digitized biomarker outcomes. For each combination of study, outcome and disease, select the data row where follow-up is closest to duration of the trial:

```{r}
d5 <- d4[
  outcome.cat == "biomarkers" & include == 1 &
    extraction.source != "graph_extracted_data" & 
    !is.na(m.e) &
    !is.na(sd.c) &
    sd.c > 0
]

d5[, grp := .GRP, by = .(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by = grp]
setorder(d5, -year)

d5[, n_studies := .N, keyby=outcome]
d6 <- d5[n_studies > 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
# X1
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1_bio_v2, fig.width=9, fig.height=3}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X) 
if(interactive())
  png("figures/test02.png", width=9, height=3, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("Outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


### Table of SMD for each non-digitized biomarker ($k = 1$)

```{r}
d6 <- d5[n_studies == 1]

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d6, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
X2 <- d6[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
```

### Forest plot of table 

```{r table_1b_bio_v2, fig.width=10, fig.height=4}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X)

if(interactive())
  png("figures/test02.png", width=10, height=4, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("Outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```


### SMD table with all studies (any $k$)

```{r}
meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e * direction, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c * direction, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year), 
                   subgroup = outcome,
                   data     = d5, 
                   sm       = "SMD")
smeta01 <- summary(meta01)
```

### Format Table 

```{r}
X1 <- with(smeta01, data.frame(outcome  = bylevs,
                               Studies = k.w,
                               SMD     = smeta01$within.random$TE,
                               seSMD   = smeta01$within.random$seTE,
                               lower   = smeta01$within.random$lower,
                               upper   = smeta01$within.random$upper,
                               pvalue  = smeta01$within.random$p,
                               I2      = I2.w))
# X1
X2 <- d5[, .(Cases = sum(n.e), Controls=sum(n.c)), by=outcome]
X <- as.data.table(merge(X1, X2, by="outcome"))
X[, CI := paste0("(", do_format(lower), 
                    " - ", do_format(upper), ")")]
X[, SMD_CI := paste(do_format(SMD), CI)]
X[, pval := format.pval(round(pvalue, 3), digits=3, eps = 1e-3)]
X[, I2_pct := format(round(I2*100))]
setorder(X, -Studies, -Cases)
X <- X[Studies > 0]
```

### Forest plot of table 

```{r table_1_suppl_bio_v2, fig.width=10, fig.height=5}
meta01 <- metagen(TE=SMD, seTE = seSMD, studlab = outcome, data = X) 
if(interactive())
  png("figures/test02.png", width=10, height=5, unit="in", res=100)
forest(meta01, digits.se = 2, digits.mean = 2, xlim=c(-2, 2), digits.pval = 3,
       overall=FALSE, overall.hetstat = FALSE,
       leftcols = c("studlab", "Studies", "Cases", "Controls", 
                    "effect.ci", "pval", "I2_pct"),
       rightcols = FALSE,
       leftlabs = c("outcome", "Studies", "Cases", "Controls",
                    "SMD [95% CI]", "p-value", "I2 %"),
       smlab = "SMD")
if(interactive()) dev.off()
```




## Appendix

**R** session info:
```{r}
sessionInfo()
```


