---
title: "PICO 3: Leave One Out Influence Analyses -- Pico 3"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.retina = 4, dev = c('png', 'pdf'))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data

```{r}
d4 <- readRDS("data_derived/pico3_4.rds")

d5 <- d4[outcome.cat == "behavioral" & include == 1]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
d5[, studlab := paste0(author, " ", year), by = .I]
setorder(d5, -year)
```


```{r}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = studlab,
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]
```


```{r}
# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]

d6[, outcome_abbrev := abbreviate(outcome, minlength = 10)]
d6[, model.cat_abbrev := abbreviate(model.cat, minlength = 5)]
```

Outcome sub-categories with more than 1 study:
```{r}
tab <- d6[, .N, outcome.sub.cat][N > 1][order(-N)]
uni_cats <- tab[, outcome.sub.cat] 
tab 
```


### Define function

```{r function_mk_leave1out}
mk_leave1out <- function(data, width=12, height=10, res=100, 
                         save_fig = interactive(), fig_path="figures/test01.png") {
  pdata <- copy(data)

  meta01 <- metacont(n.e     = n.e, 
                     mean.e  = m.e * direction, 
                     sd.e    = sd.e, 
                     n.c     = n.c, 
                     mean.c  = m.c * direction, 
                     sd.c    = sd.c, 
                     studlab = paste0(studlab, studlab_star, "; ", 
                                      outcome_abbrev, "; ", model.cat_abbrev), 
                     data    = pdata, 
                     sm      = "SMD")
  
  if(save_fig) png(fig_path, width=width, height=height, unit="in", res=res)
  forest(metainf(meta01, pooled = "random"))
  if(save_fig) dev.off()
}
```


## Leave-one-out analyses - Behavioral (|SMD| < 2) - All data

```{r, eval=FALSE}
print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", 
        "```{r leave1out_", i, ", fig.width=9, fig.height=3}\n", 
        "plotdata <- d6[outcome.sub.cat == uni_cats[", i, "]]\n",
        "mk_leave1out", "(plotdata, width = 9, height = 3)\n",
        "```\n", sep="")
  }
}

print_code(length(uni_cats))
```

### `r uni_cats[1]`

```{r leave1out_1, fig.width=11, fig.height=11}
plotdata <- d6[outcome.sub.cat == uni_cats[1]]
mk_leave1out(plotdata, width = 11, height = 11)
```

### `r uni_cats[2]`

```{r leave1out_2, fig.width=11, fig.height=10.5}
plotdata <- d6[outcome.sub.cat == uni_cats[2]]
mk_leave1out(plotdata, width = 11, height = 10.5)
```

### `r uni_cats[3]`

```{r leave1out_3, fig.width=11, fig.height=9}
plotdata <- d6[outcome.sub.cat == uni_cats[3]]
mk_leave1out(plotdata, width = 11, height = 9)
```

### `r uni_cats[4]`

```{r leave1out_4, fig.width=11, fig.height=8}
plotdata <- d6[outcome.sub.cat == uni_cats[4]]
mk_leave1out(plotdata, width = 11, height = 8)
```

### `r uni_cats[5]`

```{r leave1out_5, fig.width=11, fig.height=6}
plotdata <- d6[outcome.sub.cat == uni_cats[5]]
mk_leave1out(plotdata, width = 11, height = 6)
```

### `r uni_cats[6]`

```{r leave1out_6, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[6]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[7]`

```{r leave1out_7, fig.width=11, fig.height=3}
plotdata <- d6[outcome.sub.cat == uni_cats[7]]
mk_leave1out(plotdata, width = 11, height = 3)
```

## Leave-one-out analyses - Behavioral (|SMD| < 2) -  Excluding graph extracted data

### Format data

```{r}
d5 <- d4[outcome.cat == "behavioral" & include == 1 & 
           extraction.source != "graph_extracted_data"]
d5[, grp := .GRP, by=.(author, year, outcome, model.cat, model.spec, intervention.type)]
d5 <- d5[, .SD[which.min(abs(duration - followup))], by=grp]
d5[, studlab := paste0(author, " ", year), by = .I]
setorder(d5, -year)
```


```{r}
meta_obj <- metacont(n.e      = n.e, 
                     mean.e   = m.e * direction, 
                     sd.e     = sd.e, 
                     n.c      = n.c, 
                     mean.c   = m.c * direction, 
                     sd.c     = sd.c, 
                     studlab  = studlab,
                     subgroup = outcome.sub.cat,
                     data     = d5,
                     sm       = "SMD") 
d5$SMD <- meta_obj$TE
d6 <- d5[abs(SMD) < 2]
```


```{r}
# Add * to study-label for duplicated studies:
d6[, studlab := paste0(author, " ", year), by = .I]
d6[, studlab_n := .N, by = .(outcome.sub.cat, studlab)]
d6[, studlab_star := fifelse(studlab_n > 1, "*", "")]

d6[, outcome_abbrev := abbreviate(outcome, minlength = 10)]
d6[, model.cat_abbrev := abbreviate(model.cat, minlength = 5)]
```

Outcome sub-categories with more than 1 study:
```{r}
tab <- d6[, .N, outcome.sub.cat][N > 1][order(-N)]
uni_cats <- tab[, outcome.sub.cat] 
tab 
```


```{r, eval=FALSE}

print_code <- function(n) {
  stopifnot(n >= 1)
  for(i in 1:n) {
    cat("\n### `r uni_cats[", i, "]`\n\n", 
        "```{r leave1out_", i, "_v2, fig.width=9, fig.height=3}\n", 
        "plotdata <- d6[outcome.sub.cat == uni_cats[", i, "]]\n",
        "mk_leave1out", "(plotdata, width = 9, height = 3)\n",
        "```\n", sep="")
  }
}

print_code(length(uni_cats))
```

### `r uni_cats[1]`

```{r leave1out_1_v2, fig.width=11, fig.height=5}
plotdata <- d6[outcome.sub.cat == uni_cats[1]]
mk_leave1out(plotdata, width = 11, height = 5)
```

### `r uni_cats[2]`

```{r leave1out_2_v2, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[2]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[3]`

```{r leave1out_3_v2, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[3]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[4]`

```{r leave1out_4_v2, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[4]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[5]`

```{r leave1out_5_v2, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[5]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[6]`

```{r leave1out_6_v2, fig.width=11, fig.height=4}
plotdata <- d6[outcome.sub.cat == uni_cats[6]]
mk_leave1out(plotdata, width = 11, height = 4)
```

### `r uni_cats[7]`

```{r leave1out_7_v2, fig.width=11, fig.height=3}
plotdata <- d6[outcome.sub.cat == uni_cats[7]]
mk_leave1out(plotdata, width = 11, height = 3)
```


## Leave-one-out analyses - Behavioral - Excluding graph-extracted data (no SMD restriction)

These analyses use behavioral data excluding observations extracted from graphs. No observations are excluded based on the magnitude of the SMD. Each iteration excludes one effect estimate at a time.

### Format data

```{r prepare_leave1out_no_graph_unfiltered}
d_no_graph_unfiltered <- d4[
  outcome.cat == "behavioral" &
    include == 1 &
    extraction.source != "graph_extracted_data"
]

d_no_graph_unfiltered[
  ,
  grp := .GRP,
  by = .(
    author,
    year,
    outcome,
    model.cat,
    model.spec,
    intervention.type
  )
]

d_no_graph_unfiltered <- d_no_graph_unfiltered[
  ,
  .SD[which.min(abs(duration - followup))],
  by = grp
]

setorder(d_no_graph_unfiltered, -year)
```

### Format effect-estimate labels

```{r prepare_leave1out_no_graph_unfiltered_labels}
d_no_graph_unfiltered[
  ,
  studlab := paste0(author, " ", year)
]

d_no_graph_unfiltered[
  ,
  studlab_star := fifelse(.N > 1, "*", ""),
  by = .(outcome.sub.cat, studlab)
]

d_no_graph_unfiltered[
  ,
  outcome_abbrev := abbreviate(
    outcome,
    minlength = 10
  )
]

d_no_graph_unfiltered[
  ,
  model.cat_abbrev := abbreviate(
    model.cat,
    minlength = 5
  )
]
```

Behavioral outcome subcategories represented by more than one effect estimate:

```{r list_leave1out_no_graph_unfiltered_categories}
tab_no_graph_unfiltered <- d_no_graph_unfiltered[
  ,
  .N,
  by = outcome.sub.cat
][N > 1][order(-N)]

uni_cats_no_graph_unfiltered <-
  tab_no_graph_unfiltered$outcome.sub.cat

tab_no_graph_unfiltered
```

### Leave-one-effect-out plots

```{r leave1out_no_graph_unfiltered, results='asis', fig.width=11, fig.height=8}
stopifnot(length(uni_cats_no_graph_unfiltered) >= 1)

for (i in seq_along(uni_cats_no_graph_unfiltered)) {
  current_category <- uni_cats_no_graph_unfiltered[i]

  cat(
    "\n\n### ",
    current_category,
    "\n\n",
    sep = ""
  )

  plotdata_no_graph_unfiltered <- d_no_graph_unfiltered[
    outcome.sub.cat == current_category
  ]

  mk_leave1out(
    plotdata_no_graph_unfiltered,
    width = 11,
    height = 8,
    save_fig = FALSE
  )
}
```

## Leave-one-out analyses - Behavioral - All data (no SMD restriction)

These analyses use all included behavioral data. No observations are excluded based on extraction source or the magnitude of the SMD. Each iteration excludes one effect estimate at a time.

### Format data

```{r prepare_leave1out_all_unfiltered}
d_all_unfiltered <- d4[
  outcome.cat == "behavioral" &
    include == 1
]

d_all_unfiltered[
  ,
  grp := .GRP,
  by = .(
    author,
    year,
    outcome,
    model.cat,
    model.spec,
    intervention.type
  )
]

d_all_unfiltered <- d_all_unfiltered[
  ,
  .SD[which.min(abs(duration - followup))],
  by = grp
]

setorder(d_all_unfiltered, -year)
```

### Format effect-estimate labels

```{r prepare_leave1out_all_unfiltered_labels}
d_all_unfiltered[
  ,
  studlab := paste0(author, " ", year)
]

d_all_unfiltered[
  ,
  studlab_star := fifelse(.N > 1, "*", ""),
  by = .(outcome.sub.cat, studlab)
]

d_all_unfiltered[
  ,
  outcome_abbrev := abbreviate(
    outcome,
    minlength = 10
  )
]

d_all_unfiltered[
  ,
  model.cat_abbrev := abbreviate(
    model.cat,
    minlength = 5
  )
]
```

Behavioral outcome subcategories represented by more than one effect estimate:

```{r list_leave1out_all_unfiltered_categories}
tab_all_unfiltered <- d_all_unfiltered[
  ,
  .N,
  by = outcome.sub.cat
][N > 1][order(-N)]

uni_cats_all_unfiltered <-
  tab_all_unfiltered$outcome.sub.cat

tab_all_unfiltered
```

### Leave-one-effect-out plots

```{r leave1out_all_unfiltered, results='asis', fig.width=11, fig.height=8}
stopifnot(length(uni_cats_all_unfiltered) >= 1)

for (i in seq_along(uni_cats_all_unfiltered)) {
  current_category <- uni_cats_all_unfiltered[i]

  cat(
    "\n\n### ",
    current_category,
    "\n\n",
    sep = ""
  )

  plotdata_all_unfiltered <- d_all_unfiltered[
    outcome.sub.cat == current_category
  ]

  mk_leave1out(
    plotdata_all_unfiltered,
    width = 11,
    height = 8,
    save_fig = FALSE
  )
}
```
## Appendix

**R** session info:
```{r}
sessionInfo()
```

