---
title: "PICO 3: Individual level raw data analyses"
author: "Rune Haubo B Christensen & Sabine Færge"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
## Context

For more than 20 studies in PICO 3, individual-level data were received directly from the study authors. These raw data are not publicly available in this repository.

The individual-level data were analysed separately, and the resulting summary statistics were incorporated into the shared PICO 3 dataset used in the meta-analyses. Consequently, this file cannot be reproduced without access to the original individual-level data, but it is retained in the repository to document the analytical workflow.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Load packages:
```{r, warning=FALSE, message=FALSE}
library(readxl)
library(meta)
library(grid)
library(ggplot2)
library(data.table)
```

Read functions: 
```{r, include_functions}
source("functions/meta_analysis_functions.R")
source("functions/misc_functions.R")
source("functions/format_functions.R")
```


## Read data
Read data and create ASCII-compatible outcome labels for plotting and abbreviation.

```{r}
FILE <- "data_ipd/Restructured collection of individual-level data PICO 3.xlsx"
stopifnot(file.exists(FILE))
sheets <- excel_sheets(FILE)

# Read all sheets:
dt <- lapply(sheets, function(sheet) quick_read_excel(FILE, sheet = sheet)) |> 
  rbindlist(fill = TRUE)
dt[, study_id := paste(author, year)]

# Convert non-ASCII to ASCII:
dt[, outcome_orig := outcome]
dt[, outcome := stringi::stri_enc_toascii(outcome_orig)]

# Abbreviate outcome:
dt[, outcome_abbrev := as.factor(abbreviate(outcome, minlength = 10))]

# Check that outcome_abbrev is a 1-to-1 mapping of outcome:
is_1to1 <- function(dt, col1, col2) {
  dt <- as.data.table(dt)
  n_pairs <- uniqueN(dt, by = c(col1, col2))
  n1 <- uniqueN(dt[[col1]])
  n2 <- uniqueN(dt[[col2]])
  n_pairs == n1 && n_pairs == n2
}
stopifnot(is_1to1(dt, "outcome", "outcome_abbrev"))
```


## Data summaries

Non-ASCII cases of `outcome`:
```{r}
dt[outcome != outcome_orig, .N, .(study_id, outcome, outcome_orig)]
```

Missing values:
```{r}
show_na(dt)
dt[!is.na(model.spec)] |> print(topn = 2)
dt[, .N, keyby = model.spec] |> totpct()
```


### 1-way tables

```{r}
dt[, .N, keyby = study_id] |> totpct()
dt[, .N, keyby = group] |> totpct()
dt[, .N, keyby = dosis] |> totpct()
dt[, .(no_studies = uniqueN(study_id), N=.N), keyby = outcome] |> 
  totpct(n_col = "N")
```


### Dose

```{r}
ggplot(dt, aes(dosis)) + geom_histogram(bins = 30)
```

Dose for control group is always zero.
```{r}
stopifnot(dt[group == "control", all(dosis == 0)])
```

Distribution of dose for intervention group:
```{r}
ggplot(dt[group != "control"], aes(dosis)) + geom_histogram(bins = 30) + 
  scale_x_log10() + 
  facet_wrap(~group)
```


## Outcome value

Numerical summaries:
```{r}
dt[, as.list(summary(value)), group]
```

Cases where value is zero:
```{r}
dt[value == 0]
```

### Illustrations

A histogram simply shows the huge differences in scale:
```{r}
ggplot(dt, aes(value)) + geom_histogram(bins = 30)
```

Excluding cases where `value == 0` the histogram on the log-scale is bimodal:
```{r}
ggplot(dt[value > 0], aes(value)) + geom_histogram(bins = 30) + 
  scale_x_log10()
```

By study ID we don't see any bell-shaped distributions. 
```{r, fig.height=10}
ggplot(dt, aes(value, fill = group)) + geom_histogram(bins = 30) + 
  theme(legend.position = "bottom") + 
  facet_wrap(~ study_id, scales = "free_x") + 
  xlab("")
```

Some distributions are bimodal indicating different scales for different outcomes:
```{r, fig.height=10}
ggplot(dt, aes(value, fill = group)) + geom_histogram(bins = 30) + 
  theme(legend.position = "bottom") + 
  facet_wrap(study_id ~ outcome_abbrev, scales = "free_x") + 
  xlab("")
```

On the log-scale:
```{r, fig.height=10}
ggplot(dt[value > 0], aes(value, fill = group)) + geom_histogram(bins = 30) + 
  theme(legend.position = "bottom") + 
  facet_wrap(study_id ~ outcome_abbrev, scales = "free_x") + 
  scale_x_log10() + 
  xlab("") + ylab("")
```


## Overview forest plots

Table of N & mean(SD) by study and outcome:
```{r}
tab <- dt[, .(n = .N, 
              m = mean(value), 
              sd = sd(value)), 
          keyby = .(author, year, study_id, outcome_abbrev, outcome, group)]

tab <- dcast(tab, study_id + author + year + outcome + outcome_abbrev ~ 
               group, 
             value.var = c("n", "m", "sd"))
setnames(tab, function(nm) gsub("_control", ".c", nm, fixed = TRUE))
setnames(tab, function(nm) gsub("_intervention", ".e", nm, fixed = TRUE))
nm <- names(tab)
tab[, .SD, .SDcols = grep("study_id|outcome_abbrev|\\.e$|\\.c$", nm)] |>
  print(digits = 3)
```

```{r, mk_plot_fun, echo=FALSE}
mk_plot <- function(meta_obj, width=12, height=10, res=100, 
                    save_fig = interactive(), fig_path="figures/test01.png", ...) {
  if(save_fig) png(fig_path, width=width, height=height, unit="in", res=res)
  forest(meta_obj, 
         digits.sd = 2, 
         digits.mean = 2, 
         xlim=c(-2, 2),
         print.subgroup.name = FALSE,
         overall = FALSE, 
         overall.hetstat = FALSE,
         common = FALSE,
         random = FALSE,
         resid.hetstat = FALSE,
         test.overall.common = FALSE,
         test.overall.random = FALSE)
         
  if(save_fig) dev.off()
  invisible(NULL)
}
```


```{r, warning=FALSE}
meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c, 
                   sd.c     = sd.c, 
                   studlab  = paste0(study_id, ", ", outcome_abbrev), 
                   data     = tab,
                   sm       = "SMD") 
```

```{r ow_forestplot01, fig.width=11, fig.height=8}
mk_plot(meta01, width=11, height=8)
```


## Comparison with supplied summary data
The supplied study-level summary records corresponding to the available IPD
are marked `include == 0`, because the IPD-derived summary statistics were
used in their place. The following plot compares the two sources.

```{r}
d <- readRDS("data_derived/pico3_1.rds")
d3 <- readRDS("data_derived/pico3_3.rds")
d4 <- readRDS("data_derived/pico3_4.rds")
```

```{r, fig.width=11, fig.height=7}
d5 <- merge(d4[include == 0], tab[, .(author, year, outcome, outcome_abbrev)],
            by = c("author", "year", "outcome"))

meta01 <- metacont(n.e      = n.e, 
                   mean.e   = m.e, 
                   sd.e     = sd.e, 
                   n.c      = n.c, 
                   mean.c   = m.c, 
                   sd.c     = sd.c, 
                   studlab  = paste0(author, " ", year, ", ", outcome_abbrev), 
                   data     = d5,
                   sm       = "SMD") 
mk_plot(meta01, width=11, height=9)
```


## Appendix

### Functions

```{r, mk_plot_fun, echo=TRUE}
```


### **R** session info

```{r}
sessionInfo()
```







